An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects\\ in Time-Frequency Representations

### 2020-07-24 Notes/results.tex
matlab
xp

To quantify the global quality (\ie~not depending on $\ell$) of the forecasting approaches, we evaluate the Experimental Mean Square Error $\mathrm{MSE_{xp}}(\tilde\bx)$ of the forward forecast extended signals, namely:
\begin{align}
\label{eq:mse}
\mathrm{MSE_{\xp}}(\tilde\bx) &= \dfrac1{L}\|\tilde\bx -\bx^\mathrm{ext}\|^2 \\[-1mm]
\nonumber
&=\! \dfrac1{L}\sum_{\ell=1}^L \bmu_{\xp}[N\!-\!1\!+\!\ell]^2 \!+\! \bgamma_{\xp}[N\!-\!1\!+\!\ell,N\!-\!1\!+\!\ell] .
\end{align}
where $\bx^\mathrm{ext}$ is the ground-truth extended signal, that is: $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$. Then, as long as the bias $\bmu[N-1+\ell]$ and the variance $\bgamma[N-1+\ell,N-1+\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either. Corresponding results are given in Table~\ref{tab:mse.sine}. They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension. Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data. Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.

\begin{table}
\centering
\caption{AHM Signal: Performance of the Extension Methods}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
   \cline{2-3} & Mean & Standard deviation & \\
   \hhline{|=#=|=|=|}
   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.152$ \\
   \hline
   Symmetric & $1.019\times 10^{1}$ & $1.192\times 10^{2}$ & $0.002$ \\
   \hline
   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.537$\\
   \hline
   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.331$ \\
   \hline
   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.120$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}

@@

From that long signal, we build a dataset of $378$ non\-overlapping signals of 60 seconds, \ie~$N=6000$. On each of these pieces of signal, we implement the forecasting method introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}. However, the TBATS extension method is not implemented here because of its excessive computing time. We forecast $7$-second-long extensions on each segment of the signal, corresponding to $L =700$. Thus, in order to catch slowly varying dynamical behaviors, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$. As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$. The resulting MSE~\eqref{eq:mse} averaged with respect to the simulations are given in Table~\ref{tab:THO}. The MSE of {\sf SigExt} is higher than the MSE of the other methods. This is mainly caused by the presence of a few segments of the whole signal involving complicated behaviors, that are unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}. The left of Fig.~\ref{fig:THO.failure} illustrates one of those cases, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal. The EDMD and Symmetric extensions are more robust to those situations, as shown in this example, and in Table~\ref{tab:THO}. Nevertheless, {\sf SigExt} provides a sufficiently relevant extension to give TF representations sparingly affected by boundary effects. On the right of Fig.~\ref{fig:THO.failure}, we display a comparison between the right boundary of SST of the same segment of signal (top-right), and its boundary-free SST obtained after the {\sf SigExt} forecasting (bottom-right). The extension of the instantaneous frequency visible on the right side of the image illustrates the reduction of boundary effects produced despite an inaccurate signal forecasting.

>>>>

To quantify the global quality (\ie~not depending on $\ell$) of the forecasting approaches, we evaluate the Experimental Mean Square Error $\mathrm{MSE_{xp}}(\tilde\bx)$ of the forward forecast extended signals, namely:
\begin{align}
\label{eq:mse}
\mathrm{MSE_{\xp}}(\tilde\bx) &= \dfrac1{L}\|\tilde\bx -\bx^\mathrm{ext}\|^2 \\[-1mm]
\nonumber
&=\! \dfrac1{L}\sum_{\ell=1}^L \bmu_{\xp}[N\!-\!1\!+\!\ell]^2 \!+\! \bgamma_{\xp}[N\!-\!1\!+\!\ell,N\!-\!1\!+\!\ell] .
\end{align}
where $\bx^\mathrm{ext}$ is the ground-truth extended signal, that is: $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$. Then, as long as the bias $\bmu[N-1+\ell]$ and the variance $\bgamma[N-1+\ell,N-1+\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either. The numerical results when $L=1000$ are given in Table~\ref{tab:mse.sine}. They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension. Besides, even though the other more sophisticated methods, like GPR, are able to give slightly better forecasts, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data. Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.

\begin{table}
\centering
\caption{AHM Signal: Performance of the Extension Methods}
\begin{tabular}{|c||c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE averaged on $50$ realizations}  & \multirow{2}{80pt}{\centering Averaged computing time (sec.)} \\
    &  & \\
   \hhline{|=#=|=|}
   {\sf SigExt} & $1.458$ & $0.152$ \\
   \hline
   Symmetric & $8.735$ & $0.002$ \\
   \hline
   EDMD & $2.313$ &$2.537$\\
   \hline
   GPR & $1.441$ &$146.331$ \\
   \hline
   TBATS & $1.501$ & $1267.120$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}

@@

From that long signal, we build a dataset of $379$ non\-overlapping signals of 60 seconds, \ie~$N=6000$. On each of these pieces of signal, we implement the forecasting method introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}. However, the TBATS extension method is not implemented here because of its excessive computing time. We forecast $7$-second-long extensions on each segment of the signal, corresponding to $L =700$. Thus, in order to catch slowly varying dynamical behaviors, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$. As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$. The resulting MSE~\eqref{eq:mse} averaged with respect to the simulations are given in Table~\ref{tab:THO}. The MSE of {\sf SigExt} is higher than the MSE of the other methods. This is mainly caused by the presence of a few segments of the whole signal involving complicated behaviors, that are unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}. The left of Fig.~\ref{fig:THO.failure} illustrates one of those cases, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal. The EDMD and Symmetric extensions are more robust to those situations, as shown in this example, and in Table~\ref{tab:THO}. Nevertheless, {\sf SigExt} provides a sufficiently relevant extension to give TF representations sparingly affected by boundary effects. On the right of Fig.~\ref{fig:THO.failure}, we display a comparison between the right boundary of SST of the same segment of signal (top-right), and its boundary-free SST obtained after the {\sf SigExt} forecasting (bottom-right). The extension of the instantaneous frequency visible on the right side of the image illustrates the reduction of boundary effects produced despite an inaccurate signal forecasting.
